---
title: "Introductory Statistics Examples"
subtitle: "04 - Simple Linear Regression"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    highlight: tango
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
IntroAnalysis:::ma223_setup()
library(learnr)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.cap = "Exercise")
```

```{css, echo=FALSE}
.tipbox {
  padding: 1em 1em 1em 1em;
  border: 1px solid black;
  border-radius: 10px;
  box-shadow: 2px 2px #888888;
}

.keyidea {
  border: 1px solid #54585A;
  background: #4F758B;
  color: white;
}

.tip {
  background: lightyellow;
}

.alert {
  background: #E87722;
  color: white;
}

.title{
  color: #800000;
  background: none;
}

.subtitle{
  color: #696969;
  font-style: normal;
}

.author{
  font-weight: normal;
  font-style: italic;
}

h2{
  color: white;
  background-color: #800000;
}

h3{
  color: #800000;
  border-bottom: thick solid #696969;
}

h4{
  color: #800000;
}

blockquote {
  font-size: inherit;
  padding: 0px 20px;
}

div.alert p a {
  color: #DFD1A7;
}
```



## Background
When purchasing or selling a home, your real estate agent helps you determine if a house is priced correctly.  This involves comparing the price of the home with its "comparables" (similar homes in the area) to determine if the price is similar.  One way of doing this is to build a statistical model which predicts the sale price of a home based on various characteristics.  We have data on over 500 sales from real estate transactions across the United States.  Each record corresponds to the sale of a particular home and includes the following variables:

  - `ID`: Unique identification for each sale.
  - `Price`: Sale price of home (\$).
  - `Area`: Finished area of residence (square feet).
  - `BedRooms`: Number of bedrooms in the residence.
  - `BathRooms`: Number of bathrooms in the residence.
  - `AC`: Indicator of whether home has central air conditioning (1 = Yes, 0 = No).
  - `Garage`: Number of cars that garage will hold.
  - `Pool`: Indicator of whether home has a swimming pool (1 = Yes, 0 = No).
  - `Year`: Year property was originally constructed.
  - `Quality`: Indicator of the quality of the construction (1 = High quality, 2 = Medium, 3 = Low).
  - `Style`: Indicator of the architectural style (mapping unknown).
  - `Lot`: Lot size (square feet).
  - `Highway`: Indicator of whether home is located near highway (1 = Yes, 0 = No).

The data is available in `R` (`RealEstate`).



## Questions
Suppose a couple is considering an addition onto their home. They would like to estimate how much they should invest in such an addition.

### Exercise: Assessing the Question
> There are two ways a person might approach this objective:
>
>  - How much can they expect the value of their home to increase given the addition will increase the square footage of their home by $x$ square feet?
   - Given that after the addition, their home will have a square-footage of $x$, how much do they expect their home to be worth on the market?
>
> Compare and contrast these two questions.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=d9143a01-353a-4210-a92f-ac800129d506&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Graphical Summary
> Construct a graphic summarizing the relationship between the size of a home (its square footage) and the sale price of the home.  What does the graphic communicate (keeping in mind the objective)?

```{r s4-ex-graph, exercise=TRUE, exercise.eval=FALSE}
qplot() +
  labs()
```

```{r s4-ex-graph-solution}
qplot(data = RealEstate,
      y = `Price`,
      x = `Area`,
      geom = "point") +
  labs(y = "Sale Price of Home ($)",
       x = "Living Area (sq-ft)")
```

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=7e43222b-ac7e-46d4-b434-ac80012a9bd3&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Correlation
> Which of the following _most likely_ represents the correlation coefficient relating the sale price to the living area of the home?
>
>  -  -159
   -  -0.82
   -  -0.21
   -  0.21
   -  0.82
   -  159

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=96ffead2-be12-47fc-b6ec-ac80012b2c4d&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Model for Data Generating Process
> Using mathematical notation, write the model for the data generating process viewing the sale price as a linear function of the living area of the home.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=684c36e0-f374-41e2-9ef1-ac80012bd1dc&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Role of Error Term
> State one source of variability in the sale price being captured by the error term in the above model.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=354aa4b8-5523-457b-b198-ac80012c9492&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Computing Least Squares Estimates
> Compute the least squares estimates for the parameters in the model specified in the previous part.  Assuming the deterministic portion of that model is correctly specified, interpret the slope in the context of this problem.

```{r s4-ex-ls, exercise=TRUE, exercise.eval=FALSE}
realestate.model = specify_mean_model()

estimate_parameters()
```

```{r s4-ex-ls-solution}
realestate.model = specify_mean_model(Price ~ 1 + Area, data = RealEstate)

estimate_parameters(realestate.model)
```
   
<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=a34d303c-273b-4229-a1d9-ac80012d5324&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Estimating Mean Response
> Based on this data, and assuming the deterministic portion of the model is correctly specified, on average, how much does a home with 2600 square feet of living area sell for?  Why should we not use the model to estimate the average sale price of a home with 6000 square feet of living area?

```{r s4-model, include=FALSE}
realestate.model = specify_mean_model(Price ~ 1 + Area, data = RealEstate)
```

```{r s4-ex-mean, exercise=TRUE, exercise.eval=FALSE, exercise.setup="s4-model"}
estimate_mean_response()
```

```{r s4-ex-mean-solution}
estimate_mean_response(realestate.model, Area = 2600)
```
   
<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=bdf95103-81b6-421d-9bb1-ac80012e41c0&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Hypotheses
> Using mathematical notation, specify the null and alternative hypotheses which capture the following research question: is there evidence that the sale price of a home is linearly related to its square footage?

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=ce8328d3-c93b-4628-aa28-ac80012f27f8&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Performing Inference with CI
> Assuming the data is consistent with the conditions of the classical regression model, compute the appropriate 95% confidence interval for each of the parameters.  Based on the analysis, what conclusions can we draw regarding the hypotheses specified in the previous part?

```{r s4-ex-ci, exercise=TRUE, exercise.eval=FALSE, exercise.setup="s4-model"}
estimate_parameters()
```

```{r s4-ex-ci-solution}
estimate_parameters(realestate.model, confidence.level = 0.95,
                    assume.identically.distributed = TRUE,
                    assume.normality = TRUE)
```
   
<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=f8247cab-0f90-408d-85df-ac8001301614&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

