---
title: "Introductory Statistics Examples"
subtitle: "02 - Sampling Distributions and Confidence Intervals"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    highlight: tango
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
IntroAnalysis:::ma223_setup()
library(learnr)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.cap = "Exercise")
```

```{css, echo=FALSE}
.tipbox {
  padding: 1em 1em 1em 1em;
  border: 1px solid black;
  border-radius: 10px;
  box-shadow: 2px 2px #888888;
}

.keyidea {
  border: 1px solid #54585A;
  background: #4F758B;
  color: white;
}

.tip {
  background: lightyellow;
}

.alert {
  background: #E87722;
  color: white;
}

.title{
  color: #800000;
  background: none;
}

.subtitle{
  color: #696969;
  font-style: normal;
}

.author{
  font-weight: normal;
  font-style: italic;
}

h2{
  color: white;
  background-color: #800000;
}

h3{
  color: #800000;
  border-bottom: thick solid #696969;
}

h4{
  color: #800000;
}

blockquote {
  font-size: inherit;
  padding: 0px 20px;
}

div.alert p a {
  color: #DFD1A7;
}
```



## Background
Large surveys are commonly conducted to inform policy decisions. The Study on the Efficacy of Nosocomial Infection Control (SENIC Project) was to determine whether infection surveillance and control programs have reduced the rates of nosocomial (hospital-acquired) infection in the United States hospitals. This data set consists of a random sample of 113 hospitals selected from across the US.  Each record corresponds to information on the hospital's performance. The data contains the following variables:

  - `ID`: Unique hospital identifer.
  - `Length of Stay`: Average length of stay of all patients in hospital (days).
  - `Age`: Average age of patients (in years)
  - `Infection Risk`: estimated probability of acquiring infection in hospital (percent).
  - `Routine Culturing Ratio`: Ratio of number of cultures performed to the number of patients without
signs or symptoms of hospital-acquired infection, times 100.
  - `Routine Xray Ratio`: Ratio of number of X-rays performed to number of patients without signs or
symptoms of pneumonia, times 100.
  - `Number of Beds`: number of beds in hospital during study period.
  - `Medical School`: Indicator of whether the hospital is associated with a medical school (1 = Yes, 2 =
No).
  - `Region`: Indicator of the geographic region for hospital (1 = NE, 2 = NC, 3 = S, 4 = W).
  - `Average Census`: number of patients per day in hospital during study period.
  - `Number Nurses`: number of full-time equivalent registered and licensed practical nurses
during study period (number of full time plus one half the number of part time).
  - `Available Facilities`: Percent of 35 potential facilities and services that are provided by the hospital.

The data is available in `R` (`senic`).


## Questions

### Exercise: Interpreting Sampling Distributions
> Suppose we are interested in the proportion of hospitals which are affiliated with a medical school.  Below is a model for the sampling distribution of the proportion of hospitals affiliated with a medical school.  The model is based on 5000 bootstrap replications.  Based on the model, is it reasonable that 25% of hospitals are affiliated with a medical school?  Explain.

```{r fig-a, echo=FALSE, fig.cap="Model of the sampling distribution of the sample proportion for a random sample of 113 hospitals. Model is based on 5000 bootstrap replicates."}
set.seed(123)
prop.distn <- data_frame(
  props = rbinom(5000, size = nrow(senic),
                 prob = mean(senic$`Medical School`==1)) / nrow(senic)
)

ggplot(data = prop.distn,
       mapping = aes(x = props)) +
  geom_density() +
  labs(y = "",
       x = "Sample Proportion Hospitals\nAffiliated with Medical School") +
  scale_x_continuous(breaks = seq(0.05, 0.30, 0.05)) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=b131904d-5a1b-4dca-aa4d-ac80011717fc&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

<div class = "tipbox keyidea">
  Sampling distributions are centered on the true value of the parameter while _models_ for sampling distributions are centered on the statistic from the sample.
</div>


### Exercise: Mathematical Model
> Suppose we are interested in estimating the average number of nurses staffed by a hospital.  Using mathematical notation, write the model for the data generating process which corresponds to this research objective.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=3e0eb457-d3ef-4f03-8da8-ac80011897c0&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Computing a CI
> Using the available data compute a 98% confidence interval for the average number of nurses staffed by a hospital in the US based on the available data.  You may assume the sample is representative of all hospitals in the US; you may also assume the number of nurses staffed by one hospital is independent of the number of nurses staffed by any other hospital.

```{r s2-ex-ci, exercise=TRUE, exercise.eval=FALSE}
senic.model = specify_mean_model()

estimate_parameters()
```

```{r s2-ex-ci-solution}
senic.model = specify_mean_model(`Number Nurses` ~ 1, data = senic)

estimate_parameters(senic.model, confidence.level = 0.98,
                    assume.identically.distributed = TRUE)
```

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=df4cb9e0-bf4f-47ea-b8ba-ac80011a5e02&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

<div class = "tipbox tip">
  Note that the computer _always_ assumes observations are independent of one another (and therefore, we do not need to specify this)!
</div>


### Exercise: Interpreting a CI
> Interpret the confidence interval obtained in the previous part.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=27b928e7-1bf1-4ef8-a9ee-ac80011ba626&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

<div class = "tipbox tip">
  Never attach a "% statement" with the values of the CI computed.
</div>


### Exercise: Hypotheses
> A policy adviser argues that the data provides evidence that on average, hospitals staff more than 155 nurses.  Is this a valid interpretation (assuming the conditions discussed in part (c))?  Explain.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=5aa2d3b8-e092-41cf-b7f4-ac80011d0936&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Comparing Distributions
> In the context of estimating the number of nurses staffed by hospitals, compare and contrast the distribution of the sample and the sampling distribution of the sample mean.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=06278703-5c43-499d-bfc4-ac80011eaf2d&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

