---
title: "Introductory Statistics Examples"
subtitle: "08 - Assessing Conditions in ANOVA"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    highlight: tango
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
IntroAnalysis:::ma223_setup()
library(learnr)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.cap = "Exercise")
```

```{css, echo=FALSE}
.tipbox {
  padding: 1em 1em 1em 1em;
  border: 1px solid black;
  border-radius: 10px;
  box-shadow: 2px 2px #888888;
}

.keyidea {
  border: 1px solid #54585A;
  background: #4F758B;
  color: white;
}

.tip {
  background: lightyellow;
}

.alert {
  background: #E87722;
  color: white;
}

.title{
  color: #800000;
  background: none;
}

.subtitle{
  color: #696969;
  font-style: normal;
}

.author{
  font-weight: normal;
  font-style: italic;
}

h2{
  color: white;
  background-color: #800000;
}

h3{
  color: #800000;
  border-bottom: thick solid #696969;
}

h4{
  color: #800000;
}

blockquote {
  font-size: inherit;
  padding: 0px 20px;
}

div.alert p a {
  color: #DFD1A7;
}
```



## Background
Physical therapy is a vital part of the recovery process for any surgery, particularly surgery which corrects motor function, such as knee or hip surgery.  A rehabilitation center researcher was interested in examining the relationship between physical fitness prior to surgery of persons undergoing corrective knee surgery and time required in physical therapy until successful rehabilitation.  Patient records in the rehabilitation center were examined, and 24 male subjects ranging in age from 18 to 30 years who had undergone similar corrective knee surgery during the past year were selected for the study.  The number of days required for successful completion of physical therapy and the prior physical fitness status (below average, average, above average) for each patient was obtained. 

The data contains the following variables:

  - `Subject`: unique patient identifer.
  - `Prior Activity Level`: indicator of the activity level of the subject prior to surgery.
  - `Recovery Time`: time (days) required to complete physical therapy successfully.
  - `Age`: age (years, rounded to nearest tenth) of patient.

The data is available in `R` (`Therapy`).



## Questions
Researchers are primarily interested in determining if, on average, the recovery time (in days) following surgery differs for subjects classified with different activity levels prior to their surgery.  As a result, we have been studying the following model for the data generating process:

$$
\begin{aligned}
  (\text{Recovery Time})_i &= \mu_1 (\text{Above Average Activity})_i + \mu_2 (\text{Average Activity})_i \\
    &\qquad + \mu_3 (\text{Below Average Activity})_i + \epsilon_i
\end{aligned}
$$

where

$$
\begin{aligned}
  (\text{Above Average Activity})_i &= \begin{cases}
    1 & \text{if i-th patient has above average physical activity prior to surgery} \\
    0 & \text{otherwise}
    \end{cases} \\
  (\text{Average Activity})_i &= \begin{cases}
    1 & \text{if i-th patient has average physical activity prior to surgery} \\
    0 & \text{otherwise}
    \end{cases} \\
  (\text{Below Average Activity})_i &= \begin{cases}
    1 & \text{if i-th patient has below average physical activity prior to surgery} \\
    0 & \text{otherwise}
    \end{cases}
\end{aligned}
$$
The data is summarized below.

```{r, echo=FALSE}
ggplot(data = Therapy) +
  aes(y = `Recovery Time`,
      x = `Prior Activity Level`) +
  labs(y = "Recovery Time",
       x = "Prior Activity Level") +
  geom_boxplot()
```


This model is fit using the following code.

```{r s8-model, echo=TRUE}
therapy.model = specify_mean_model(`Recovery Time` ~ `Prior Activity Level`,
                                   data = Therapy)
```




### Exercise: Compute Residuals
> Obtain the residauls and fitted values for the above model, and store them in a dataset called `therapy.diag`.  What does the residual for each observation tell you?

```{r s8-ex-resid, exercise=TRUE, exercise.eval=FALSE, exercise.setup="s8-model"}
therapy.diag = obtain_diagnostics()
```

```{r s8-ex-resid-solution, exercise.setup="s8-model"}
therapy.diag = obtain_diagnostics(therapy.model)
```

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=f61987d7-e477-43e4-b7f5-ac80014c9d9e&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

<div class = "tipbox tip">
You must always obtain the residuals from the full model (not the one under the null hypothesis) before trying to assess conditions.
</div>


### Exercise: Assessing Normality
> Is it reasonable to assume the error in the above model comes from a Normal distribution?  Explain. 

```{r s8-ex-normality, exercise=TRUE, exercise.eval=FALSE, exercise.setup="s8-ex-resid-solution"}
ggplot() +
  aes() +
  labs()
```

```{r s8-ex-normality-solution}
ggplot(data = therapy.diag) +
  aes(sample = .resid) +
  labs(y = "Sample Quantiles",
       x = "Theoretical Quantiles") +
  geom_qq()
```

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=5ba96623-6777-4f74-b562-ac80014d5968&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Assessing Homoskedasticity
> Is it reasonable that the variability in the error of the recovery time is the same for each of the prior physical activity levels?  Explain. 

```{r s8-ex-homoskedasticity, exercise=TRUE, exercise.eval=FALSE, exercise.setup="s8-ex-resid-solution"}
ggplot() +
  aes() +
  labs()
```

```{r s8-ex-homoskedasticity-solution}
ggplot(data = therapy.diag) +
  aes(y = .resid,
      x = `Prior Activity Level`) +
  labs(y = "Residuals",
       x = "Prior Activity Level") +
  geom_boxplot()
```

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=4407bccb-c01c-43a2-8312-ac80014dc552&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

<div class = "tipbox tip">
There is more than one way to assess constant variance in the ANOVA setting:

- We can use the residuals against the fitted values as in regression.
- We can look at the raw data.
- We can use the residuals against the factor levels (groups).
</div>


### Exercise: Assessing Independence
> Is it reasonable to assume the error in the recovery time for one individual is independent of the error in the recovery time for any other individual?  Explain.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=747b5bb4-9574-4f7e-a13f-ac80014eab9c&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>


### Exercise: Assessing Mean 0?
> Explain why we do not need to worry about whether the deterministic portion of the model is correctly specified.

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=3060b5a1-c1a7-4429-9540-ac80014f5dd6&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>



### Exercise: Compute a P-Value
> Given your conclusions thus far, compute a p-value for addressing the question of interest on which we can rely.

```{r s8-ex-pval, exercise=TRUE, exercise.eval=FALSE, exercise.setup="s8-ex-resid-solution"}
therapy.model.h0 = specify_mean_model()

compare_models()
```

```{r s8-ex-pval-solution}
therapy.model.h0 = specify_mean_model(`Recovery Time` ~ 1, data = Therapy)

compare_models(therapy.model,
               therapy.model.h0,
               assume.constant.variance = TRUE,
               assume.normality = TRUE)
```

<div class = "tipbox alert">
<iframe src="https://rose-hulman.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=fb04fd66-555f-4fc3-a43f-ac80015032fb&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>
</div>

